Towards Comparable Active Learning

### 2023-06-02 doc/al_benchmark_paper.tex
batch bald 50

\title{Towards Comparable Active Learning}
\author{%
	Thorben Werner 
	\thanks{Institute of Computer Science - Information Systems and Machine Learning Lab (ISMLL)} 
	\\
	University of Hildesheim\\
	Universitätsplatz 1\, 31141 Hildesheim \\
	\texttt{werner@ismll.de} \\
	% examples of more authors
	\And
	Johannes Burchert$^*$ \\
	University of Hildesheim\\
	Universitätsplatz 1, 31141 Hildesheim \\
	\texttt{burchert@ismll.de} \\
	\AND
	Prof. Lars Schmidt-Thieme$^*$ \\
	University of Hildesheim\\
	Universitätsplatz 1, 31141 Hildesheim \\
	\texttt{schmidt-thieme@ismll.uni-hildesheim.de} \\
	% \And
	% Coauthor \\
	% Affiliation \\
	% Address \\
	% \texttt{email} \\
	% \And
	% Coauthor \\
	% Affiliation \\
	% Address \\
	% \texttt{email} \\
}

@@

\begin{abstract}
	%Active Learning (AL) is a technique used in machine learning to improve data efficiency by selecting the most informative samples for annotation. 
	In this paper we address the issue of inconsistent results in active learning (AL) literature.
	Previous papers are constantly reporting significant performance improvements, while subsequent literature fails to reproduce those results. This inconsistency leads to a chaotic landscape of AL algorithms.
	To the best of our knowledge, we propose the first AL benchmark that tests algorithms in 3 major domains; Tabular, Image and Text.
	Furthermore, we discuss overlooked problems for reproducing AL experiments that can lead to unfair comparisons and increased variance in the results.
	We propose a framework for choosing hyperparameters of AL setups and accurately controlling the experiments.
	Using this framework, we report empirical results for 6 algorithms on 7 datasets and aggregate them into a domain-specific ranking of AL algorithms.
\end{abstract}

\section{Introduction}
Active Learning (AL) plays an important role in our society that applies machine learning to more and more areas and therefore has a high demand for labeled data in more and more areas.
A problem that concerns academic researchers and practitioners in businesses alike and even could be extended to education in schools and hobbyists around the world.
On top of providing a principled way to labeled unlabeled datasets, active learning is one of the two major approaches besides semi-supervised learning to make deep learning models more data efficient by requiring only a limited set of manually labeled data.
Both approaches are at their core orthogonal and can freely be combined and therefore we should continue our research efforts for both approaches. \\
Among others, the authors of \cite{zhou2021towards} have pointed out severe inconsistencies in results of AL papers in recent years. 
In their supplementary materials they conducted a meta analysis of reported results of several different AL algorithms and found that all considered algorithms only provided significant lifts in their own original papers, while all following literature reported performances no better that uncertainty sampling, or in some cases no better than random sampling for the same algorithm.
The result of these inconsistencies is a chaotic landscape of AL algorithms where every paper claims to archive state-of-the-art results by significantly outperforming everyone else, while the vast majority of results proves to be non-reproducible.

\subsection{Contributions}
\begin{enumerate}
	\item Evaluation of Active Learning algorithms on datasets from 3 different domains, including synthetic data that highlights principled shortcomings of existing approaches.
	\item Novel experimental protocol for seeding the experiment with 3 different seeds to allow full control and reproducibility and analysis of how many restarts are required to converge to the true median performance reliably.
	\item Simple algorithm for an Oracle-Curve that can be constructed greedily and does not rely on search.
\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{(From Vijaya) Problem Description}
Given $n = l+u$ data points with $l \in \mathbb{N}$ many labeled examples $\mathcal{L} = \{(x_1, y_1),\ldots, (x_l,y_l)\}$, $u \in \mathbb{N}$ many unlabeled examples $\mathcal{U} = \{x_{l+1},\ldots,x_{n}\}$, a budget $\mathbb{N} \ni b \le u$ and an annotator $A: \mathbb{R}^M \to \mathbb{R}^C$ that can label $x$. We call $x \in \mathbb{R}^M$, $y \in \mathbb{R}^C$ predictors and labels respectively where $(x,y)$ are drawn from an unknown distribution $\rho$. 
Find an acquisition function $F: \mathcal{U},\mathcal{L}, b \mapsto \mathcal{U}_b$ that selects $b$ many examples from $\mathcal{U}$ such that $\mathcal{U}_b \subseteq \mathcal{U}$, $|\mathcal{U}_b|=b$ for labeling and the expected loss $\ell: \mathbb{R}^C \times \mathbb{R}^C \to \mathbb{R}$ (cross entropy loss) of the machine learning algorithm $\hat{y}$ (ex. classification) trained on $\mathcal{L} \cup \{\left(x, A(x)\right) \mid x \in \mathcal{U}_b\}$ is minimal: 
$$\min \quad \mathbb{E}_{(x,y) \sim \rho} \ell(y, \hat{y}(x))$$
Finding the optimal subset $\mathcal{U}_b$ is a combinatorial problem that is computationally not feasible for large $b$, so we allow sequential construction of the subset.
%To construct the active learning setting from a labeled dataset $\mathcal{D} = \{(x_1, y_1),\ldots, (x_n,y_n)\}$, we first split it into $\mathcal{D}_\text{train}$, $\mathcal{D}_\text{val}$ and $\mathcal{D}_\text{test}$ and then suppress the labels of $\mathcal{D}_\text{train}$ to form the unlabeled pool $\mathcal{U}$ and form and initial labeled pool by uniformly sampling $k$ number of instances per class, removing them from $\mathcal{U}$ and recovering their label $\mathcal{L} := \{(x_1, A(x_1)),\ldots, (x_{k*C},A(x_{k*C}))\} \quad s.t. \quad \mathcal{U} \cap \mathcal{L} = \emptyset$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{(From Lars) Problem Description}
Given two spaces $\mathcal{X}:=\mathcal{R}^M$ and $\mathcal{Y}:=\mathcal{R}^C$, a sample $\mathcal{D}_1,\ldots,\mathcal{D}_N \subseteq (\mathcal{X}\times \mathcal{Y})^*$ of sequences of pairs $(x,y)$  from an unknown distribution $p$ called datasets and a number $B\in\mathcal{N}$ with $B < |\mathcal{D}|$. \\
Given two functions $\ell:\mathcal{Y}\times \mathcal{Y}\rightarrow \mathcal{R}$ called loss, and $A: (\mathcal{X} \times \mathcal{Y})^* \times \mathcal{X}^* \rightarrow \mathcal{Y}^{\mathcal{X}}$ called learning algorithm, find a function
\begin{align*}
	a: (\mathcal{X}\times \mathcal{Y})^* \times \mathcal{X}^* \rightarrow \{0,1\}^*
\end{align*}
called acquisition function, s.t. the expected loss of a model learned on all predictors plus $B$ acquired targets is minimal.
%\begin{align*}
%	\mathbb{E}_{\mathcal{D}\text{train},\mathcal{D}\text{test}\sim p}   &
%	% \frac{1}{|\mathcal{D}\text{test}|} \sum_{(x,y)\in\mathcal{D}\text{test}}
%	% \operatorname{avg} \{
%	%  \ell(y, \hat y(x)) \mid (x,y)\in\mathcal{D}\text{test} \}
%	\operatorname{avg}\limits_{(x,y)\in\mathcal{D}\text{test}}
%	\ell(y, \hat y(x)) 
%	\\
%	\text{with }
%	\hat y:= & A( (\mathcal{D}\text{train}_{n_1},\ldots,\mathcal{D}\text{train}_{n_B}), \mathcal{D}\text{train}|_{\mathcal{X}})
%	\\ 
%	n_b := & \text{index}( a( (\mathcal{D}\text{train}_{n_1},\ldots,\mathcal{D}\text{train}_{n_{b-1}}), \mathcal{D}\text{train}|_{\mathcal{X}}) ),
%	\quad b\in 1{:}B
%\end{align*}
%(\textit{We would need to switch from lowest expected loss to highest AUC}) \\
Even though the acquisition function in principle could output the full subset of $\mathcal{D}_\text{train}$, the combinatorial problem is computationally not feasible for large $B$ and we allow sequential construction of the subset. \\
%
%To construct the active learning setting from existing labeled datasets, we first split it into $\mathcal{D}_\text{train}$, $\mathcal{D}_\text{val}$ and $\mathcal{D}_\text{test}$ and then suppress the labels $\mathcal{Y}$ of $\mathcal{D}_\text{train}$ to form the unlabeled pool $\mathcal{U} := \mathcal{X}^*$ and form and initial labeled pool by uniformly sampling $k$ number of instances per class from $\mathcal{U}$ and recovering their label $\mathcal{L} := (\mathcal{X}, \mathcal{Y})^{k*C}$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Overview}
We constrain our work on pool-based active learning where a pool of unlabeled samples is fixed at the start of each experiment and samples are chosen sequentially.
Specifically, we are not experimenting on so-called batch active learning, where at each iteration multiple unlabeled samples are chosen at the same time.
Even though batch AL is the more active research domain, it does not have a principled advantage over single-sample AL except speed of computation.
Not only is the problem of optimizing a portfolio of unlabeled samples more complicated to solve, the algorithms also have systematically less information per sample to work with.
For this reason we propose to focus more research effort on single-sample AL to find better algorithms in an environment that is simpler to solve and easier to control.
A performance comparison of batch AL and single-sample AL can be found in Fig. \ref{fig:ablation_bald}, which reproduces the message of Figure 1 from the paper \cite{gal2017deep} that proposed BALD, one of the SOTA algorithms for AL.
We can see that the batched version of BALD \cite{krizhevsky2009learning} can at most perform on-par with the single-sample algorithm.
Fig. \ref{fig:ablation_bald} also serves as a proof of concept for our provided code base.
Table \ref{tab:benchmark_comparison} shows a feature comparison between our proposed benchmark and several existing benchmarks in the literature, as well as methodological AL papers with experiments on at least two data domains.
\begin{table}[h]
	\centering
	\begin{tabular}{l | c c c c c}
		Paper & Sampling & \#Datasets & \#Domains & \#Algorithms & Oracle \\
		\hline
		Beck et al. \cite{beck2021effective} & batch & 4 & 1 & 7 & - \\
		Hu et al. \cite{hu2021towards} & batch & 5 & 2 & 13 & - \\
		Li et al. \cite{li2022empirical} & batch & 5 & 1 & 13 & - \\
		Zhou et al. \cite{zhou2021towards} & batch & 3 & 2 & 2 & \checkmark \\
		\textbf{Ours} & single & 7 & 3 & 6 & \checkmark 
	\end{tabular}
	\caption{Comparison of our benchmark with the existing literature. Oracle curves serve as an approximation of the best possible AL algorithm.}
	\label{tab:benchmark_comparison}
\end{table}

%Manual NIPS Benchmark paper eval (out of 6):
%\begin{itemize}
%	\item Has new method: 1
%	\item Multiple domains: 2
%	\item $> 5$ datasets: 3
%	\item $> 5$ baselines: 4
%\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}
{\color{red} Version: Braindump}\\
\begin{wrapfigure}{r}{0.5\textwidth}
	\centering
	\includegraphics[width=\linewidth]{img/ablation_bald}
	\caption{Shows a reproduction of the performance analysis of BALD from the original authors \cite{gal2017deep} in comparison to entropy sampling and an adaptation of BALD for batch AL \cite{kirsch2019batchbald} with a batch size of 40.}
	\label{fig:ablation_bald}
\end{wrapfigure}
Many different algorithms have been proposed for active learning. 
In this work we focus on those approaches that have shown consistent results over the years as well as some of the new approaches.
AL algorithms can be categorized into two classes: Geometric approaches and uncertainty-based approaches.
Geometric approaches use clustering techniques to partition the data and then sample their unlabeled points based on the clusters.
They often use the current classification model $\hat y_i$ to encode the data into a latent space to improve the performance of their clustering.
This benchmark includes the following geometric approaches: CoreSet \cite{sener2017active}, BADGE \cite{ashdeep} and TypiClust \cite{hacohen2022active}.
Uncertainty-based approaches use metrics to measure the classifiers state.
Commonly, a proxy for the sought after uncertainty of the model for a given datapoint is the distance of that point to the various decision boundaries, measured via the softmax output of the model.
This benchmark includes Shannon-Entropy sampling \cite{wang2014new}, margin sampling \cite{wang2014new} and BALD \cite{kirsch2019batchbald} \\ [1mm]
%
Some previous work also aimed to provide a benchmark suite for active learning:
The authors of \cite{beck2021effective} and \cite{li2022empirical} both focus on active learning in the image domain.
While \cite{beck2021effective} discuss a new metric to measure AL performance, which they call ``Label Efficiency'' and provide experiments on many common configurations of data preparation, model training and other hyperparameters, \cite{li2022empirical} focuses on combined approaches of AL and semi-supervised learning to aid model training.
The authors of \cite{hu2021towards} study models that are learned with AL techniques in the image and text domain.
They test for several different properties of the models including robustness, response to compression techniques and final performance.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Evaluation}\label{sec:evaluation}
Following \cite{zhou2021towards}, the quality of an active learning algorithm is evaluated by an "anytime" protocol that incorporates classification performance at every iteration, not just the final performance after the budget is exhausted.
We employ the normalized area under the accuracy curve (AUC):
\begin{equation}\label{eq:auc}
	\operatorname{AUC}(\mathcal{D}_{test}, \hat y, B) := \frac{1}{B} \sum_{i=1}^{B} \operatorname{Acc}(\mathcal{D}_{test}, \hat y_i)
\end{equation}
where $\hat y_i$ is the (re-)trained classification model after the i-\textit{th} iteration.
To mimic the leave-one-out protocol for cross-validation we will restart each experiment multiple times.
Each restart will retain the train/test split (often given by the dataset itself), but introduces a new validation split.
The AUC incorporates performance in early stages (low budget) as well as capabilities to push the classifier in later stages (high budget).
A good AL algorithm should be able to perform well in both scenarios. 
Since AL has high variance and is prone to outliers, we propose to report median AUC instead of mean AUC values. \\ [1mm]
The AUC is dependent on the chosen budget, so we need a general rule on how to set this hyperparameter that does not inherently benefit a subset of algorithms.
In this work, we choose the budget per dataset to be the first point at which any algorithm (except oracle) manages to reach a percentage of the upper bound performance measured on the full dataset.
Even though we would like to propose a single percentage value for all datasets, we found that different data modalities and use cases need different percentages to produce sensible budgets.
We propose the following values: \textbf{Tabular}: 99\%, \textbf{Image}: 90\% and \textbf{Text}: 95\%. \\ [1mm]
Additionally, we provide evidence in Fig. \ref{fig:restarts} that previous works have not evaluated their experiments with a sufficient number of restarts.
To create Fig. \ref{fig:restarts} we used 50 restarts from the margin/random sampling algorithm on the splice dataset.
From these 50 runs we uniformly sampled subsets of runs and calculated the median AUC for this subset.
One of these median AUC values corresponds to one cross-validated experiment sampled from the distribution of experiments that are restarted exactly this many times.
To create one slice in Fig. \ref{fig:restarts}, we drew 50 samples from this distribution.
Each box-plot represents the variance of an evaluation if conducted with the respective number of restarts.
We can observe that low repetitions ($<10$) provide an uncertain evaluation where lucky and unlucky draws of the same experiment give drastically different median AUC values.
\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{img/ablation_restarts.png}
	\caption{Random draws from an experimental distribution on the Splice datasets with different numbers of repetitions. Each point on the Y-axis represents a cross-validated result that could have been reported in a paper. This analysis shows the drastic differences in performance one could observe even when repeating an experiment 2-10 times.}
	\label{fig:restarts}
\end{figure}
To reliably arrive at the true median AUC, we propose to repeat every experiment 50 times, as only $>42$ repitions don't produce outliers anymore (as indicated by the rightmost columns in Fig \ref{fig:restarts}).
One way to reduce the number of necessary repetitions is to reduce the amount of variance in the experiment through specialized seeding (discussed in the next section).
We ultimately decided in favor of high variance and high number of repetitions as the high variance accurately reflects real world applications of AL.

>>>>

\newcommand{\R}{\mathbb{R}}
\newcommand{\EE}{\mathbb{E}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\Y}{\mathcal{Y}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\LL}{\mathcal{L}}
\newcommand{\test}{\text{test}}
\newcommand{\train}{\text{train}}

\title{Towards Comparable Active Learning}
\author{%
	Anonymous Authors
	%Thorben Werner 
	%\thanks{Institute of Computer Science - Information Systems and Machine Learning Lab (ISMLL)} 
	%\\
	%University of Hildesheim\\
	%Universitätsplatz 1\, 31141 Hildesheim \\
	%\texttt{werner@ismll.de} \\
	% examples of more authors
	%\And
	%Johannes Burchert$^*$ \\
	%University of Hildesheim\\
	%Universitätsplatz 1, 31141 Hildesheim \\
	%\texttt{burchert@ismll.de} \\
	%\AND
	%Prof. Lars Schmidt-Thieme$^*$ \\
	%University of Hildesheim\\
	%Universitätsplatz 1, 31141 Hildesheim \\
	%\texttt{schmidt-thieme@ismll.uni-hildesheim.de}
}

@@

\begin{abstract}
	Active Learning has received significant attention in the field of machine learning for its potential in selecting the most informative samples for labeling, thereby reducing data annotation costs. However, the lack of reproducibility of performance gains reported in recent literature has created a chaotic landscape in Active Learning research. 
	This paper addresses these issues of inconsistent results in active learning (AL) literature.
	%Previous papers are constantly reporting significant performance improvements, while subsequent literature fails to reproduce those results. This inconsistency leads to a chaotic landscape of AL algorithms.
	To the best of our knowledge, we propose the first AL benchmark that tests algorithms in 3 major domains: Tabular, Image and Text.
	Furthermore, we highlight overlooked problems for reproducing AL experiments that can lead to unfair comparisons and increased variance in the results.
	To tackle these challenges, we propose an experimental protocol to accurately control the experiments.
	We report empirical results for 6 widely used algorithms on 7 datasets and aggregate them into a domain-specific ranking of AL algorithms.
\end{abstract}

\section{Introduction}
Deep neural networks (NN) have produced state-of-the-art results on many important supervised learning tasks.
Since Deep NNs usually require large amounts of labeled training data, Active Learning (AL) can be used to instead to select the most informative samples out of a large pool of unlabeled data, so that only these samples need to be labeled.
It has been shown that a small labeled set of this nature can be used to train well-performing models \cite{beck2021effective, hu2021towards, li2022empirical, zhou2021towards}. \\
%On top of providing a principled way to labeled unlabeled datasets, active learning is one of the two major approaches besides semi-supervised learning to make deep learning models more data efficient by requiring only a limited set of manually labeled data.
%Both approaches are at their core orthogonal and can freely be combined and therefore we should continue our research efforts for both approaches. \\
In the last decade many different algorithms for AL have been proposed.
Even though, almost every method has reported lifts of their approach over random selection and all predecessors \footnote{out of all considered algorithms for this paper, only BALD \cite{gal2017deep} did not establish a new SOTA performance in their results sections}, AL research faces two central difficulties:
(i) The experiments are often carried out on different datasets and model architectures, hindering direct comparison,
(ii) The reported results prove to be very difficult to reproduce.
While multiple benchmark suites have been proposed to solve (i), to the best of our knowledge, we are the first to compare AL algorithms on all 3 data domains of vector, image and text.
Regarding (ii), \cite{zhou2021towards} has pointed out severe inconsistencies in results of AL papers in recent years. 
They conducted a meta analysis of reported results of several different AL algorithms and found that all considered algorithms only provided significant lifts in their own original papers, while following literature reported performances no better that uncertainty sampling, or in some cases no better than random sampling for the same algorithm (\cite{zhou2021towards} Appendix A).
The result of these inconsistencies is a chaotic landscape of AL algorithms where every paper claims to archive state-of-the-art results by significantly outperforming everyone else, while the vast majority of results proves to be non-reproducible. \\
In this work we propose an evaluation protocol that was designed to cope with high variance in the performances of AL algorithms as well as being fully controllable regardless of the combination of dataset, model and AL algorithm.\\
We focus our work on single-sample pool-based AL where a pool of unlabeled samples is fixed at the start of each experiment and samples are chosen sequentially.
Specifically, we are not experimenting on so-called Batch Active Learning, where at each iteration multiple unlabeled samples are chosen at the same time.
Even though most proposed algorithms (and benchmarks) only do Batch AL, it does not have a principled advantage over single-sample AL except speed of computation.
The problem of optimizing a portfolio of unlabeled samples more complicated to solve and the algorithms also have systematically less information per sample to work with.
%For this reason we propose to focus more research effort on single-sample AL to find better algorithms in an environment that is simpler to solve and easier to control.
A performance comparison of batch AL and single-sample AL can be found in Fig. \ref{fig:ablation_bald}.
We can see that for two established algorithms BALD \cite{krizhevsky2009learning} and BADGE \cite{ashdeep} Batch AL can at most perform on-par with single-sample AL. \\
Table \ref{tab:benchmark_comparison} shows a feature comparison between our proposed benchmark and several existing benchmarks in the literature.
The table also includes methodological AL papers with experiments on at least two data domains.
\begin{table}
	\centering
	\begin{tabular}{l | c c c c c}
		Paper & Sampling & \#Datasets & \#Domains & \#Algorithms & Oracle \\
		\hline
		Beck et al. \cite{beck2021effective} & batch & 4 & 1 & 7 & - \\
		Hu et al. \cite{hu2021towards} & batch & 5 & 2 & 13 & - \\
		Li et al. \cite{li2022empirical} & batch & 5 & 1 & 13 & - \\
		Zhou et al. \cite{zhou2021towards} & batch & 3 & 2 & 2 & \checkmark \\
		\textbf{Ours} & single & 7 & 3 & 6 & \checkmark 
	\end{tabular}
	\caption{Comparison of our benchmark with the existing literature. Oracle curves serve as an approximation of the best possible AL algorithm.}
	\label{tab:benchmark_comparison}
\end{table}

\subsection*{Contributions}
\begin{enumerate}
	\item Evaluation of Active Learning algorithms on datasets from 3 different domains, including synthetic data that highlights principled shortcomings of existing approaches.
	\item Novel experimental protocol for seeding the experiment with 3 different seeds to allow full control, reproducibility and analysis of how many restarts are required to converge to the true median performance reliably.
	\item Simple algorithm for an Oracle that can be constructed greedily and does not rely on search.
\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Problem Description}
Given two spaces $\mathcal{X}, \mathcal{Y}$, $n=l+u$ data points with $l \in \mathbb{N}$ labeled examples $\mathcal{L} = \{(x_1, y_1),\ldots, (x_l,y_l)\}$, $u \in \mathbb{N}$ unlabeled examples $\mathcal{U} = \{x_{l+1},\ldots,x_{n}\}$, a budget $\mathbb{N} \ni b \le u$ and an annotator $A: \mathbb{R}^\mathcal{X} \to \mathbb{R}^\mathcal{Y}$ that can label $x$. 
We call $x \in \mathcal{X}$, $y \in \mathcal{Y}$ predictors and labels respectively where $(x,y)$ are drawn from an unknown distribution $\rho$. 
Find an acquisition function $a: \U^{(i)},\LL^{(i)} \mapsto x^{(i)}$ that iteratively selects the next unlabeled point $x^{(i)} \in \U^{(i)}$ for labeling
\begin{align*}
	\LL^{(i+1)} &\gets \LL^{(i)} \cup \{\left(x^{(i)}, A(x^{(i)})\right)\} \\
	\U^{(i+1)} &\gets \U^{(i)} \setminus x^{(i)} %, i \in 1{:}B
\end{align*}
so that the expected loss $\ell: \mathbb{R}^\mathcal{Y} \times \mathbb{R}^\mathcal{Y} \to \mathbb{R}$ of the machine learning algorithm $\hat y$ after $B$ iterations is minimal: 
$$\min \quad \mathbb{E}_{(x,y) \sim \rho} \ell(y, \hat{y}(\LL^{(B)}))$$
This formulation is a special case of the problem formulation in Appendix \ref{app:problem_formulation}.

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}
\begin{figure}
	\centering
	\includegraphics[width=0.49\linewidth]{img/ablation_bald}
	\includegraphics[width=0.49\linewidth]{img/ablation_bald}
	\caption{Shows a reproduction of the performance analysis of BALD from the original authors \cite{gal2017deep} in comparison to entropy sampling and an adaptation of BALD for batch AL \cite{kirsch2019batchbald} with a batch size of 40.}
	\label{fig:ablation_bald}
\end{figure}
%Many different algorithms have been proposed for Active Learning. 
%In this work we focus on those approaches that have shown consistent results over the years as well as some of the new approaches.
%AL algorithms can be categorized into two classes: Geometric approaches and uncertainty-based approaches.
%Geometric approaches use clustering techniques to partition the data and then sample their unlabeled points based on the clusters.
%They often use the current classification model $\hat y_i$ to encode the data into a latent space to improve the performance of their clustering.
%This benchmark includes the following geometric approaches: CoreSet \cite{sener2017active}, BADGE \cite{ashdeep} and TypiClust \cite{hacohen2022active}.
%Uncertainty-based approaches use metrics to measure the classifiers state.
%Commonly, a proxy for the sought after uncertainty of the model for a given datapoint is the distance of that point to the various decision boundaries, measured via the softmax output of the model.
%This benchmark includes Shannon-Entropy sampling \cite{wang2014new}, margin sampling \cite{wang2014new} and BALD \cite{kirsch2019batchbald} \\ [1mm]
%
Multiple benchmark suites have already been proposed for Active Learning:
The authors of \cite{beck2021effective} and \cite{li2022empirical} both focus exclusively on Batch AL in the image domain.
While \cite{beck2021effective} discuss a new metric to measure AL performance, which they call ``Label Efficiency'' and provide experiments on many common configurations of data preparation, model training and other hyperparameters, \cite{li2022empirical} focuses on combined approaches of AL and semi-supervised learning.
The authors of \cite{hu2021towards} study models that are trained with AL techniques in the image and text domain.
They test for several different properties of the models including robustness, response to compression techniques and final performance.
\cite{zhou2021towards} proposed an oracle algorithm for AL that uses Simulated Annealing search to approximate a solution for the optimal subset of labeled data.
Additionally, they study the generalization behavior of subsets of labeled data in the text an image domain.
The employed AL algorithms for our experiments are introduced in Section \ref{sec:sampling_strategies}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Evaluation}\label{sec:evaluation}
Following \cite{zhou2021towards}, the quality of an AL algorithm is evaluated by an ``anytime" protocol that incorporates classification performance at every iteration, as opposed to evaluating final performance after the budget is exhausted.
We employ the normalized area under the accuracy curve (AUC):
\begin{equation}\label{eq:auc}
	\operatorname{AUC}(\mathcal{D}_{test}, \hat y, B) := \frac{1}{B} \sum_{i=1}^{B} \operatorname{Acc}(\mathcal{D}_{test}, \hat y_i)
\end{equation}
where $\hat y_i$ is the (re-)trained classification model after the i-\textit{th} iteration.
To mimic the leave-one-out protocol for cross-validation we will restart each experiment multiple times.
Each restart will retain the train/test split (often given by the dataset itself), but introduces a new validation split.
The AUC incorporates performance in early stages (low budget) as well as capabilities to push the classifier in later stages (high budget).
AL algorithms have to perform well in both scenarios. 
Since AL performance has high variance and is prone to outliers, we propose to aggregate AUC values with the median instead of mean. \\ [1mm]
Since the AUC depends on the chosen budget, we need a general rule on how to set this hyperparameter that does not inherently benefit a subset of algorithms.
In this work, we choose the budget per dataset to be the first point at which any algorithm (except oracle) manages to reach a percentage of the upper bound performance measured on the full dataset.
Even though we would like to propose a single percentage value for all datasets, we found that different data modalities and use cases need different percentages to produce sensible budgets
\footnote{Since we only consider single-sample AL, increasing the budget comes at a high computational cost. Therefore we could not freely increase the budget above 2000.}.
We propose the following values: \textbf{Tabular}: 99\%, \textbf{Image}: 90\% and \textbf{Text}: 95\%. \\ [1mm]
Additionally, we provide evidence in Fig. \ref{fig:restarts} that previous works have not evaluated their experiments with a sufficient number of restarts.
To create Fig. \ref{fig:restarts} we used 50 restarts from the margin/random sampling algorithm on the splice dataset.
From these 50 runs we uniformly sampled subsets of runs and calculated the median AUC for this subset.
One of these median AUC values corresponds to one cross-validated experiment sampled from the distribution of experiments that are restarted exactly this many times.
To create one slice in Fig. \ref{fig:restarts}, we drew 50 samples from this distribution.
Each box-plot represents the variance of an evaluation if conducted with the respective number of restarts.
We can observe that low repetitions ($<10$) provide an uncertain evaluation where lucky and unlucky draws of the same experiment give drastically different median AUC values.
\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{img/ablation_restarts}
	\caption{Random draws from an experimental distribution on the Splice dataset with different numbers of repetitions. Each point on the Y-axis represents a cross-validated result that could have been reported in a paper. This analysis shows the drastic differences in performance one could observe even when repeating an experiment 2-10 times.}
	\label{fig:restarts}
\end{figure}
To reliably arrive at the true median AUC, we propose to repeat every experiment 50 times, as only $>42$ repetitions don't produce outliers anymore (as indicated by the rightmost columns in Fig \ref{fig:restarts}).
One way to reduce the number of necessary repetitions is to reduce the amount of variance in the experiment through specialized seeding (discussed in the next section).
We ultimately decided in favor of high variance and high number of repetitions as the high variance accurately reflects real world applications of AL.
Fixing for example the classifier initialization to random draws from a constant seed does not have a corresponding use-case in the real world.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Reproducibility}\label{sec:reproducibility}
A big focus in this work is to provide an experimental setup that is fully reproducible independent of the dataset, classification model or AL algorithm used.
For a fair comparison of two AL algorithms, both algorithms should receive equal starting conditions in terms of train / validation split, initialization of classifier and even the state of minor systems like the optimizer or mini-batch sampler.
Additionally, previous works have noted a significant adverse effect of training stochasticity on the performance of AL algorithms (\cite{zhou2021towards, lowell2018practical}).
Even though implementations might have their own solution to some of these problems, to the best of our knowledge no previous work has discussed this topic in detail.
At the core of this problem is the seeding utility in PyTorch, Tensorflow and other frameworks, whose default choice is a single global seed.
Since many systems draw random numbers from this seed, all of them influence each other to a point where a single additional draw can completely change the model initialization or data split, depending on the ordering of these two in the implementation.
Even though you could find workarounds like re-setting the seed multiple times, this problem also extends to every AL iteration and the systems within.
We propose an implementation that creates a separate Random Number Generator (RNG) for each of these systems to ensure equal testing conditions even when the AL algorithm, dataset or classifier changes.
We hypothesize that the insufficient setup with global seeds contributes to the on-going problem of inconsistent results of AL algorithms in different papers. \\ [1mm]
In summary, we introduce three different seeds: $s_\Omega$ for the AL algorithm, $s_\mathcal{D}$ for dataset splitting and mini batch sampling and $s_\theta$ for model initialization and sampling of dropout masks.
Unless stated otherwise, we will keep $s_\Omega$ fixed for restarts of the same experiment, while $s_\mathcal{D}$ and $s_\theta$ are incremented by 1 between restarts to introduce stochasticity into our framework.
Some algorithms require a subsample to be drawn from $\mathcal{U}$ in order to reduce the computational cost in each iteration, while others need access to the full unlabeled pool (i.e. for effective clustering).
If a subsample is required, it will be drawn from $s_\Omega$ and therefore will not influence other systems in the experiments.
For each algorithm, we decided if subsampling is required based on our available hardware, but decided against setting a fixed time limit per experiment, since this would introduce unnecessary complexity into the benchmark.
An overview of selected hyperparameters per AL algorithm can be found in Appendix \ref{app:agent_hyperparameters}.
